---
title: "Client Growth"
author: "jonathan littel"
date: "August 18, 2016"
output: 
  ioslides_presentation
---

```{r  setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE)
wd <- 'C:/Box Sync/jlittel/comms/client financials/data'
setwd(wd)
load('client_fin2.Rdata')
options(digits = 0, scipen = 12)
```

## data

- Social & environmental metrics
- Client Financials (prior and while active clients)
- Client characteristics (eg region, sector)
- Revenue, Loss (expected/realized); assumptions for operating and interest expense

`r dim(clients)[1]` observations of clients, with `r sum(clients$active_year)` observations with a loan outstanding
## Median sales by first loan size

```{r}
sales.plot
```

## Payments to producers by loan type

```{r }
purchases.plot
```

## Sales where Root was the first lender

```{r }
first.plot
```

## Clusters


```{r}
cagrs.plot
```

## Africa - Sales growth by client

```{r}
afr.plot
```
Each line is a client (for which we lent and have data for at least 3 years)



## Possible Cuts
- Agroprocessors
- Region
- Country
- Sales size at start
- additionality
- impact score

## Example target statements

For a segment, say African private enterprises with $500k - $1M in sales, working with Root enable the businesses to double their revenue in x years, increasing payments to producers by y. The cost of this was z in the first year, decreasing to i in the x year.

## Clients which started with high growth

```{r  setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE)
wd <- 'C:/Box Sync/jlittel/comms/client financials/data'
setwd(wd)
load('clientfin3.RData')
options(digits = 0, scipen = 12)
```

```{r}
client.growth.plot
```
** grey line means growth was undefined at some point - only one year, infinite **

```{r}
loan.n.rev.plot5
```
