---
title: "Summary of Client Financial data"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
wd <- 'C:/Box Sync/jlittel/comms/client financials'
setwd(wd)
load('client_fin.Rdata')
options(digits = 0, scipen = 12)
```

## The data

- Financials for ~1400 of ~2000 loans
- Additional data from SKI
- Financials for ~`r sum(!is.na(fin$sales[fin$active==TRUE]), na.rm=TRUE) / sum(fin$active) * 100`% of years with a loan outstanding
- `r n_clients` clients
- `r n_account_years` years where a client had a balance outstanding



## For clients with data

```{r echo = FALSE}
options(digits = 2)
rownames(bal_w_data) <- ''
t(bal_w_data)
options(digits = 1)
```



$`r bal_w_data[1,1] / 1e9`b in sales, on `r sum(!is.na(fin$sales[fin$active==TRUE]), na.rm=TRUE) / sum(fin$active) * 100`% of the years with a loan outstanding



## Growth rates

```{r }
(t(growth_summary))
```

Median growth: __`r growth_summary['growth_median'] * 100`%__

(from year before first loan to last year with available data)

Median CAGR: __`r (growth_summary['cagr_median']) * 100`%__

## Data is skewed

Client | Year | Sales
------------- | -------------
Inka Gold | 2012 | $1m
Inka Gold | 2013 |  $3m
Inka Gold | 2014 |  ?
Inka Gold | 2015 |  ?
Inka Gold | 2016 |  ?


If clients are growing, skew is towards lower sales


## Absolute growth


The year before the loan, clients had __$`r prettyNum(sum(cagrs$sales[cagrs$Year==cagrs$year_zero])/1e9, big.mark=',')`b__ sales, and in the most recent year for which we have sales data, __$`r prettyNum(sum(cagrs$sales[cagrs$Year==cagrs$max_sales_year])/1e9, big.mark=',')`b__.


Root helped create at least __$`r (sum(cagrs$sales[cagrs$Year==cagrs$max_sales_year]) - sum(cagrs$sales[cagrs$Year==cagrs$year_zero])) / 1e6`m__ in sales.

## Attribution

Clients where Root was the majority financer in the first year^1^ grew at a cagr of `r median(cagrs$sales_cagr[cagrs$attribution==1], na.rm=TRUE) * 100`%

^1^ Loan amount more than 60% of sales.


## Imputing sales data

We can get to a higher number for cumulative client revenue by estimating the missing data. 

The theory (and data) indicates that clients are growing, and most of the missing data is in later years. 

If we fill in the years as a root client with the sales for the most recent year we have data, cumulative sales is $`r sum(impute$sales[impute$balance>0], na.rm=TRUE) / 1e9`b. 

(from $`r bal_w_data$sum_sales / 1e9`b before imputing)

## Obligatory inscrutable graph

```{r, echo=FALSE}
library(ggplot2)
grow_graph2 
```

`r prop.table(table(cagrs$grew))[2] * 100`% of clients grew in our data.
This is likely biased low, as the data is skewed towards earlier in client's time with Root.

## Growth by #  of Years as Root Client

```{r, echo=FALSE, warning=FALSE}
yr.gr
```
Growth tapers over multiple years as clients. 